A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-11-25 kobakian-cook.Rmd
caption for figure 5

@@

The algorithm is implemented in the `sugarbag` package for `R`, named after the common name for the *Trigona carbonaria* bee – a species which builds flat layers of hexagonal brood cells, spiralling out from a central point [@PH]. This unusual pattern also provided the inspiration for the algorithm, where the maps are constructed by building out from multiple focal points on the geographic map base in a spiral fashion.

@@

The `sugarbag` package contains a set of functions that help R users to create a hexagon tile map. The algorithm presented in the `sugarbag` package operates on a set of simple feature geometry objects , known as `sf` objects[@sf]. This package allows R users to create `sf` objects by importing polygons stored in various formats. Users should provide a set of polygons that define geographic units by their administrative boundaries. The functions arrange the geographic units in order of proximity to a set of locations provided, such as the centre of major cities. The centroid location of each geographic unit is used to measure the proximity. It emphasises the major cities as population hubs, rather than emphasizing the size of large, rural geographic units.

@@

and the development version can be installed from the GitHub repository:

@@

The set of SA2 polygons for Tasmania, using the 2011 definition of SA2s, is accessed from the `absmapsdata` package. A single column of the data set is used to identify the unique areas. In this case, the unique SA2 names for each SA2 have been used.

@@

The buffer distance, hexagon size, hexagon amount to filter and width of angle are parameters that will be determined within `create_hexmap()`, if they are not provided. They are created as they are needed throughout the following example.

@@

```{r filterprocess, fig.cap = "Illustration of several steps of the algorithm: (a) ... (b) ... (c) ... (d). ", dpi = 600, echo = FALSE, height = 12}

>>>>

It is also common to treat all geographic units as equal, so that regardless of the population or other information, to generate equal size spatial polygons. The Dorling algorithm can beinduced to do this, to make equallly sized circles. @MDAC describes a group of alternative maps called "mosaic cartograms" for tile displays. The geography of the USA has been transformed into a square-binned tile map manually and made available in the R package \CRANpkg{statebins} [@statebins].The algorithm described below is an automatic layout of hexagon tiles, where a continguity constraint is relaxed. This gives equal weight to each geographic area even if the population differs.

@@

The algorithm is implemented in the \CRANpkg{sugarbag} package for `R`, named after the common name for the *Trigona carbonaria* bee – a species which builds flat layers of hexagonal brood cells, spiralling out from a central point [@PH]. This unusual pattern also provided the inspiration for the algorithm, where the maps are constructed by building out from multiple focal points on the geographic map base in a spiral fashion.

@@

The \pkg{sugarbag} package contains a set of functions that help R users to create a hexagon tile map. The algorithm presented in the \pkg{sugarbag} package operates on a set of simple feature geometry objects , known as \CRANpkg{sf} objects[@sf]. This package allows R users to create \pkg{sf} objects by importing polygons stored in various formats. Users should provide a set of polygons that define geographic units by their administrative boundaries. The functions arrange the geographic units in order of proximity to a set of locations provided, such as the centre of major cities. The centroid location of each geographic unit is used to measure the proximity. It emphasises the major cities as population hubs, rather than emphasizing the size of large, rural geographic units.

@@

\noindent and the development version can be installed from the GitHub repository:

@@

The set of SA2 polygons for Tasmania, using the 2011 definition of SA2s, is accessed from the \pkg{absmapsdata} package. A single column of the data set is used to identify the unique areas. In this case, the unique SA2 names for each SA2 have been used.

@@

The buffer distance, hexagon size, hexagon amount to filter and width of angle are parameters that will be determined within \code{create_hexmap()}, if they are not provided. They are created as they are needed throughout the following example.

@@

```{r filterprocess, fig.cap = "Illustration of key steps of the algorithm: (a) full hexagon grid is created first; (b) buffer is applied, shown as dark green circles, to accommodate irregularly shaped regions; (c, d) allocation process, relative the center of Hobart, showing the 8th centroid to be allocated. The relationship between Hobart (the cross) and the centroid (the purple triangle) is used to filter the potential locations from the grid within a wedge. Hexagons already allocated are shown in black, and the purple circle indicates the hexagon to be assigned to the 8th centroid.", dpi = 600, echo = FALSE, height = 12}

### 2020-11-25 kobakian-cook.tex
bib fixes
caption for figure 5

@@

This algorithm should be general enough to apply to other study regions
and contexts. The algorithm has only been tested using single countries.
While the buffer allows extension beyond the furthest centroids, there
is no mechanism implemented to protect the borders and ensure centroids
are placed within the geographic borders of a country. The algorithm is
also currently limited to representing each geographic area with only
one hexagon. Future developments may address these limitations, and
refine the filter steps for final selection of a hexagon grid point for
each centroid. A direct angle is currently implemented, but may narrowly
miss appropriate hexagons, a logarithmic function may help to choose a
closer hexagon to the original centroid location. This algorithm is an
effective start to creating hexagon tile maps for many geographic units.
This contributes to an extensive body of work that encourages the use of
alternative map displays for statistical displays of spatial data.

>>>>

It is also common to treat all geographic units as equal, so that
regardless of the population or other information, to generate equal
size spatial polygons. The Dorling algorithm can beinduced to do this,
to make equallly sized circles. \citet{MDAC} describes a group of
alternative maps called ``mosaic cartograms'' for tile displays. The
geography of the USA has been transformed into a square-binned tile map
manually and made available in the R package \CRANpkg{statebins}
\citep{statebins}.The algorithm described below is an automatic layout
of hexagon tiles, where a continguity constraint is relaxed. This gives
equal weight to each geographic area even if the population differs.

@@

This algorithm should be general enough to apply to other study regions
and contexts. The algorithm has only been tested using single countries.
While the buffer allows extension beyond the furthest centroids, there
is no mechanism implemented to protect the borders and ensure centroids
are placed within the geographic borders of a country. The algorithm is
also currently limited to representing each geographic area with only
one hexagon. Future developments may address these limitations, and
refine the filter steps for final selection of a hexagon grid point for
each centroid. A direct angle is currently implemented, but may narrowly
miss appropriate hexagons, a logarithmic function may help to choose a
closer hexagon to the original centroid location. This algorithm is an
effective start to automatically creating hexagon tile maps for many
geographic units. This contributes to an extensive body of work that
encourages the use of alternative map displays for statistical displays
of spatial data.
